---
title: "Chapter 3: Technologies"
---

> *Chapter goals:*
>
> 1.  *Understand the basic concept of SNP chip genotyping (also called beadchip, bead array, microarray), and the high level data handling process.*
>
> 2.  *Gain introductory knowledge of Next Generation Sequencing (NGS), and implementation of this technology to sequence DNA.*

## 3.1 SNP Array Genotyping {#sec-section1}

> This video by Dr. Mészáros is focused on understanding the SNP chip technology: how SNP chips work and why they are useful.

**Title**: SNP Chips (Introduction to genomics theory)

**Presenter(s)**: Gábor Mészáros, PhD (Institute of Livestock Sciences (NUWI), University of Natural Resources and Life Sciences)

**Length**: 28:05

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=PeFOjLEJkrA")
```

Link to video transcript [here](chapter3.1_transcript.qmd).

Powerpoint slides for this video are available [here](https://github.com/GenomicsBootCamp/IntroductionToGenomics).

------------------------------------------------------------------------

## 3.2 Next Generation Sequencing {#sec-section2}

> This section gives an introduction to Next Generation Sequencing. The first video by Dr. Kiel gives a very brief introduction to the concept of DNA sequencing, and how nucleotide bases are "read" from the DNA. The second video from ClevaLab gives a more detailed view of Next Generation Sequencing, how it differs from Sanger sequencing, and the different steps involved in sequencing a DNA sample with NGS methods, including library preparation, Sequencing by Synthesis and the sequencing reaction, cluster generation from fragments, indexing, filtering, mapping, and demultiplexing the reads.

### *How to Sequence the Human Genome*

**Title**: How to sequence the human genome

**Presenter(s)**: Mark J. Kiel, MD, PhD (Genomenon)

**Length**: 5:04

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=MvuYATh7Y74")
```

Link to video transcript [here](chapter3.2_transcript.qmd#sec-video1).

------------------------------------------------------------------------

### *Next Generation Sequencing: A Step-by-Step Guide to DNA Sequencing*

**Title**: Next Generation Sequencing: A Step-by-Step Guide to DNA Sequencing

**Presenter(s)**: ClevaLab

**Length**: 7:37

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=WKAUtJQ69n8")
```

Link to video transcript [here](chapter3.2_transcript.qmd#sec-video2).
