---
title: "Chapter 9: Advanced Topics"
---

> *Chapter goals:*
>
> 1.  *Gain a more extensive understanding of advanced genetic analyses.*
> 2.  *Understand what a copy number variant is and methods for defining CNVs in your data.*
> 3.  *Understand the concept behind Mendelian Randomization and methods for running MR.*
> 4.  *Gain a comprehensive understanding of Genomic Structural Equation Modeling (Genomic SEM) and methods for performing Genomic SEM.*
> 5.  *Understand gene x environment interactions and how they are measured.*
> 6.  *Gain understanding of Twin studies and how a twin-based analysis is run.*
> 7.  *Understand how drug target analyses and association tests are run on genetic data, and how this can help with the development of therapeutics for various disorders.*

# 9.1 Copy Number Variation {#sec-section1}

> This video by Dr. Howrigan explains copy number variation: what is a copy number variant, and how is it detected in genetic data. Additionally, examples of CNV analyses demonstrate what a CNV file format looks like, as well as output from CNV analyses, and how to perform CNV burden and association testing on that data.

**Title**: How to run Copy Number Variation (CNV) analysis

**Presenter(s)**: Daniel Howrigan, Broad Institute

**Level**: Intermediate

**Length**: 20:03

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=Lg2k2BEQIy4")
```

Link to video transcript [here](chapter9.1_transcript.qmd).

------------------------------------------------------------------------

# 9.2 Mendelian Randomization {#sec-section2}

> This video from Dr. Smith gives a very brief and simple explanation of Mendelian Randomization, and how it is an important tool in detecting causality.

**Title**: A two minute primer on Mendelian Randomization

**Presenter(s)**: George Davey Smith, TARG Bristol

**Level**: Beginner

**Length**: 2:16

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=LoTgfGotaQ4")
```

Link to video transcript [here](chapter9.2_transcript.qmd).

------------------------------------------------------------------------

# 9.3 Genomic Structural Equation Modeling {#sec-section3}

> The following six videos by Drs. Grotzinger and Nivard give a comprehensive look at Genomic Structural Equation Modeling (Genomic SEM) for examining shared and distinct genetic relationships across multiple disorders. The first two videos give an introduction to the concept and basic usage of Genomic SEM, while the latter four videos dive deeper into the methods and analyses that can be performed with this tool.

**Title**: Genomic Structural Equation Modeling: A Brief Introduction

**Presenter(s)**: Andrew Grotzinger

**Level**: Beginner

**Length**: 10:43

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=FV7tFV1p7JQ&list=PL-A34BVyxWtV2q5COofdmLJAntQKbiKYW")
```

Link to video transcript [here](chapter9.3_transcript.qmd#sec-video1).

------------------------------------------------------------------------

**Title**: Short Primer on Structural Equation Modeling (SEM) in Lavaan

**Presenter(s)**: Andrew Grotzinger

**Level**: Intermediate

**Length**: 11:34

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=eulJrhGTBgc&list=PL-A34BVyxWtV2q5COofdmLJAntQKbiKYW&index=3")
```

Link to video transcript [here](chapter9.3_transcript.qmd#sec-video2).

------------------------------------------------------------------------

**Title**: GenomicSEM: Input/Explaining how S and V are estimated and what they are

**Presenter(s)**: Michel Nivard

**Level**: Intermediate

**Length**: 23:18

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=35NrsTiTpFg&list=PL-A34BVyxWtV2q5COofdmLJAntQKbiKYW&index=4")
```

Link to video transcript [here](chapter9.3_transcript.qmd#sec-video3).

------------------------------------------------------------------------

**Title**: Working through examples on the Genomic SEM wiki one by one: munge, ldsc, usermodel functions

**Presenter(s)**: Michel Nivard

**Level**: Intermediate

**Length**: 22:54

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=Paf2JfWES6I&list=PL-A34BVyxWtV2q5COofdmLJAntQKbiKYW&index=5")
```

Link to video transcript [here](chapter9.3_transcript.qmd#sec-video4).

Link to tutorial GitHub [here](https://github.com/GenomicSEM/GenomicSEM) and Wiki [here](https://github.com/GenomicSEM/GenomicSEM/wiki).

------------------------------------------------------------------------

**Title**: Multivariate GWAS in Genomic SEM

**Presenter(s)**: Andrew Grotzinger

**Level**: Intermediate

**Length**: 30:08

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=KreSR2CMims&list=PL-A34BVyxWtV2q5COofdmLJAntQKbiKYW&index=6")
```

Link to video transcript [here](chapter9.3_transcript.qmd#sec-video5).

------------------------------------------------------------------------

**Title**: Using Genomic SEM to Understand Psychiatric Comorbidity

**Presenter(s)**: Andrew Grotzinger

**Level**: Intermediate

**Length**: 1:01:02

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=2Qyg7toEEDc")
```

Link to video transcript [here](chapter9.3_transcript.qmd#sec-video6).

------------------------------------------------------------------------

# 9.4 Interactions with Environmental Factors {#sec-section4}

> In this first part of Section 9.4, Dr. Lambert gives an introduction to the concept of interaction terms for statistical analyses: understanding how and when to use them, and how to interpret these terms. The first video gives a brief introduction, and the second video gives an introduction to interaction terms when using continuous variables.
>
> In the second part of this section, Dr. Westerman gives an introduction to gene-environment interactions in the context of psychiatric genetics, describing how GxE interactions can be detected in genetics data, how statistical power can impact the detection of GxE interactions, and describing the difference between additive and multiplicative interactions.

**Title**: Dummy variables: interaction terms explanation

**Presenter(s)**: Ben Lambert

**Level**: Beginner friendly

**Length**: 4:35

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=ScKL40dp8M4&ab_channel=BenLambert")
```

Link to video transcript [here](chapter9.4_transcript.qmd#sec-video1).

------------------------------------------------------------------------

**Title**: Continuous variables: interaction term interpretation

**Presenter(s)**: Ben Lambert

**Level**: Beginner friendly

**Length**: 4:53

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=HyA_Vgbc0t4&ab_channel=BenLambert")
```

Link to video transcript [here](chapter9.4_transcript.qmd#sec-video2).

------------------------------------------------------------------------

**Title**: Gene-environment interaction analysis

**Presenter(s)**: Kenny Westerman, Broad Institute

**Level**: Intermediate (content will be easier to those with a statistical genetics background)

**Length**: 42:28

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=wDxeO73q5k0&list=PLEEE2A91B09B77B4A&index=44&ab_channel=BroadInstitute")
```

Link to video transcript [here](chapter9.4_transcript.qmd#sec-video3).

------------------------------------------------------------------------

# 9.5 Family-based analysis {#sec-section5}

> In this video Dr. Maes describes univariate (one phenotype) methods of Twin modeling to determine familial genetic and environmental traits using OpenMx software.

**Title**: Univariate/MonoPhenotype Twin Modeling in OpenMx

**Description**:

**Presenter(s)**: Hermine H. Maes

**Level**: Intermediate

**Length**: 18:21

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=XPv80fzzfKw")
```

Link to video transcript [here](chapter9.5_transcript.qmd).

Link to [OpenMx software](https://openmx.ssri.psu.edu/).

------------------------------------------------------------------------

# 9.6 Therapeutic Implications {#sec-section6}

> In the below video, Dr. Whirl-Carrillo gives a comprehensive overview of pharmacogenetics methods using the example of the Pharmacogenomics Knowledge Base (PharmGKB) database, which is a large database of clinical information, drug label annotations, and curated pathways for annotating genes. This talk describes how this database was put together, and how it can be used for performing pharmacogenomics analyses and augmenting genomics analyses.

**Title**: Pharmacogenomics knowledge for personalized medicine

**Presenter(s)**: Cristina Rodriguez-Antona, Michelle Whirl-Carrillo

**Level**: Intermediate

**Length**: 44:27

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=YngN2Ie26PM&list=PLi2PVGG1ibsaHSzE3tmaaBiJ725e7RxsM&index=1")
```

Link to video transcript [here](chapter9.6_transcript.qmd).

Link to the [PharmGKB website](https://www.pharmgkb.org/).
