---
title: "Chapter 7: Ancestry-Specific Analyses and Considerations"
---

> *Chapter goals:*
>
> 1.  *Understand that ancestry is an important consideration for performing genetic analyses*
>
> 2.  *Understand how population stratification can affect GWAS results*
>
> 3.  *Understand considerations for PRS across different ancestries and how to calculate cross-ancestry PRS*
>
> 4.  *Understand cross-ancestry methods for GWAS and post-GWAS analyses*

# Chapter 7.1: Cross-Ancestry Analysis {#sec-section1}

> In this video, Dr. Duncan reviews the PGC PTSD freeze 1 trans-ancestry meta-analysis. She then discusses the differences in findings across ancestries and limitations of polygenic scoring across ancestries.

**Title**: Cross-ancestry PTSD and Polygenic Findings, and the Cross-Population SIG of the PGC

**Presenter(s)**: Laramie Duncan, PhD (Department of Psychiatry and Behavioral Sciences, Stanford University)

**Level**: Beginner friendly

**Length**: Video starts at 00:00 and goes until 13:39 (13:39 total)

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=p3m6MM5b-UU&t=148s")
```

Link to video transcript [here](chapter7_transcript.qmd#sec-video1).

\[Note: This paper is now available as [*Genome-wide Association Studies in Ancestrally Diverse Populations: Opportunities, Methods, Pitfalls, and Recommendations: https://doi.org/10.1016/j.cell.2019.08.051*\]](https://doi.org/10.1016/j.cell.2019.08.051)

------------------------------------------------------------------------

# Chapter 7.2: Ancestry-Specific PRS {#sec-section2}

> Dr. Alicia Martin discusses how to calculate polygenic risk scores and the differences in PRS accuracy across ancestries. She explains why there are discrepancies across ancestries and next steps to reduce disparities.

**Title**: Clinical use of current polygenic risk scores may exacerbate health disparities

**Presenter(s)**: Alicia Martin, PhD (The Broad Institute of MIT and Harvard)

**Level**: Beginner friendly

**Length**: Video starts at 32:39 and goes to 47:06 (15:45 total)

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=p3m6MM5b-UU&t=2321s")
```

Link to video transcript [here](chapter7_transcript.qmd#sec-video2).

------------------------------------------------------------------------

# Chapter 7.3: Local Ancestry and Admixed Populations {#sec-section3}

> Dr. Elizabeth Atkinson describes her Tractor method, which determines local ancestry for tracts in admixed individuals. This method can allow for Local Ancestry (LA) tracts to be incorporated in GWAS analysis. 

**Title**: Tractor: A framework enabling the well-calibrated genomic analysis of psychiatric traits across admixed populations

**Presenter(s)**: Elizabeth Atkinson, PhD (Department of Molecular and Human Genetics, Baylor College of Medicine)

**Level**: Beginner/Intermediate

**Length**: Video starts at 14:50 and ends at 32:38 (total 19:28)

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=p3m6MM5b-UU&t=890s")
```

Link to video transcript [here](chapter7_transcript.qmd#sec-video3).

Link to Tractor [GitHub](https://github.com/Atkinson-Lab/Tractor) and [Tutorial](https://github.com/Atkinson-Lab/Tractor-tutorial).
