---
title: "Chapter 9: Advanced Topics"
---

------------------------------------------------------------------------

# 9.1 Copy Number Variation {#sec-section1}

Goals of this section:

> 1.  

**Title**: How to run Copy Number Variation (CNV) analysis

**Description**: Explanation of the interaction term Part 1. This two-part short video series briefly describes how to use and interpret interaction terms.

**Presenter(s)**: Daniel Howrigan, Broad Institute

**Level**: Intermediate

**Length**: 20:03

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=Lg2k2BEQIy4")
```

Link to video transcript [here](chapter9.1_transcript.qmd).

------------------------------------------------------------------------

# 9.2 Mendelian Randomization {#sec-section2}

Goals of this section:

> 1.  

**Title**: A two minute primer on Mendelian Randomization

**Description**:

**Presenter(s)**: George Davey Smith, TARG Bristol

**Level**: Beginner

**Length**: 2:16

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=LoTgfGotaQ4")
```

Link to video transcript [here](chapter9.2_transcript.qmd).

------------------------------------------------------------------------

# 9.3 Genomic Structural Equation Modeling {#sec-section3}

Goals of this section:

> 1.  

## Genomic SEM Introduction

**Title**: Genomic Structural Equation Modeling: A Brief Introduction

**Description**:

**Presenter(s)**: Andrew Grotzinger

**Level**: Beginner

**Length**: 10:43

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=FV7tFV1p7JQ&list=PL-A34BVyxWtV2q5COofdmLJAntQKbiKYW")
```

Link to video transcript [here](chapter9.3_transcript.qmd#sec-video1).

------------------------------------------------------------------------

## Running Genomic SEM

**Title**: Short Primer on Structural Equation Modeling (SEM) in Lavaan

**Description**:

**Presenter(s)**: Andrew Grotzinger

**Level**: Intermediate

**Length**: 11:34

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=eulJrhGTBgc&list=PL-A34BVyxWtV2q5COofdmLJAntQKbiKYW&index=3")
```

Link to video transcript [here](chapter9.3_transcript.qmd#sec-video2).

------------------------------------------------------------------------

**Title**: GenomicSEM: Input/Explaining how S and V are estimated and what they are

**Description**:

**Presenter(s)**: Michel Nivard

**Level**: Intermediate

**Length**: 23:18

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=35NrsTiTpFg&list=PL-A34BVyxWtV2q5COofdmLJAntQKbiKYW&index=4")
```

Link to video transcript [here](chapter9.3_transcript.qmd#sec-video3).

------------------------------------------------------------------------

**Title**: Working through examples on the Genomic SEM wiki one by one: munge, ldsc, usermodel functions

**Description**:

**Presenter(s)**: Michel Nivard

**Level**: Intermediate

**Length**: 22:54

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=Paf2JfWES6I&list=PL-A34BVyxWtV2q5COofdmLJAntQKbiKYW&index=5")
```

Link to video transcript [here](chapter9.3_transcript.qmd#sec-video4).

------------------------------------------------------------------------

**Title**: Multivariate GWAS in Genomic SEM

**Description**:

**Presenter(s)**: Andrew Grotzinger

**Level**: Intermediate

**Length**: 30:08

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=KreSR2CMims&list=PL-A34BVyxWtV2q5COofdmLJAntQKbiKYW&index=6")
```

Link to video transcript [here](chapter9.3_transcript.qmd#sec-video5).

------------------------------------------------------------------------

**Title**: Using Genomic SEM to Understand Psychiatric Comorbidity

**Description**:

**Presenter(s)**: Andrew Grotzinger

**Level**: Intermediate

**Length**: 1:01:02

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=2Qyg7toEEDc")
```

Link to video transcript [here](chapter9.3_transcript.qmd#sec-video6).

------------------------------------------------------------------------

# 9.4 Interactions with Environmental Factors {#sec-section4}

## Statistics of the Interaction Term

Goals of this section:

> 1.  Understand the statistical basis of the interaction term

**Title**: Dummy variables: interaction terms explanation

**Description**: Explanation of the interaction term Part 1. This two-part short video series briefly describes how to use and interpret interaction terms.

**Presenter(s)**: Ben Lambert

**Level**: Beginner friendly

**Length**: 4:35

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=ScKL40dp8M4&ab_channel=BenLambert")
```

Link to video transcript [here](chapter9.4_transcript.qmd#sec-video1).

------------------------------------------------------------------------

**Title**: Continuous variables: interaction term interpretation

**Description**: Explanation of the interaction term Part 2. This two-part short video series briefly describes how to use and interpret interaction terms in the case of continuous variables.

**Presenter(s)**: Ben Lambert

**Level**: Beginner friendly

**Length**: 4:53

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=HyA_Vgbc0t4&ab_channel=BenLambert")
```

Link to video transcript [here](chapter9.4_transcript.qmd#sec-video2).

------------------------------------------------------------------------

## Gene environment interaction in psychiatric genetics

Goals of this section:

> 1.  Understand how gene environment interactions can be detected in statistical population genetics.
>
> 2.  Understand how statistical power impacts detecting gene by environment interactions.
>
> 3.  Understand additive vs multiplicative interactions.

**Title**: Gene-environment interaction analysis

**Description**: MPG Primer: Introduction to gene-environment interactions in the context of statistical genetics. Introduces gene environment interactions, statistics, variant prioritization, variance-QTLs, and different types of interactive effects.

**Presenter(s)**: Kenny Westerman, Broad Institute

**Level**: Intermediate (content will be easier to those with a statistical genetics background)

**Length**: 42:28

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=wDxeO73q5k0&list=PLEEE2A91B09B77B4A&index=44&ab_channel=BroadInstitute")
```

Link to video transcript [here](chapter9.4_transcript.qmd#sec-video3).

------------------------------------------------------------------------

# 9.5 Family-based analysis {#sec-section5}

Goals of this section:

> 1.  Understand

**Title**: Univariate/MonoPhenotype Twin Modeling in OpenMx

**Description**:

**Presenter(s)**: Hermine H. Maes

**Level**: Intermediate

**Length**: 18:21

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=XPv80fzzfKw")
```

Link to video transcript [here](chapter9.5_transcript.qmd).

------------------------------------------------------------------------

# 9.6 Therapeutic Implications {#sec-section6}

Goals of this section:

> 1.  

**Title**: Pharmacogenomics knowledge for personalized medicine

**Description**:

**Presenter(s)**: Cristina Rodriguez-Antona, Michelle Whirl-Carrillo

**Level**: Intermediate

**Length**: 44:27

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=YngN2Ie26PM&list=PLi2PVGG1ibsaHSzE3tmaaBiJ725e7RxsM&index=1")
```

Link to video transcript [here](chapter9.6_transcript.qmd).
