---
title: "Chapter 3: Technologies"
---

> *Chapter goals:*
>
> 1.  *Understand the basic concept of SNP chip genotyping (also called beadchip, bead array, microarray), and the high level data handling process.*
>
> 2.  *Gain introductory knowledge of Next Generation Sequencing (NGS), and implementation of this technology to sequence DNA.*

## 3.1 SNP array genotyping {#sec-section1}

> This video is focused on understanding the SNP chip technology: how SNP chips work and why are they useful.

**Title**: SNP Chips (Introduction to genomics theory)

**Presenter(s)**: Gábor Mészáros, Genomics Boot Camp

**Length**: 28:05

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=PeFOjLEJkrA")
```

Link to video transcript [here](chapter3.1_transcript.qmd).

Powerpoint slides for this video are available [here](https://github.com/GenomicsBootCamp/IntroductionToGenomics).

------------------------------------------------------------------------

## 3.2 Next Generation Sequencing {#sec-section2}

> This section gives an introduction to Next Generation Sequencing. The first video gives a very brief introduction to the concept of DNA sequencing, and how nucleotide bases are "read" from the DNA. The second video gives a more detailed view of Next Generation Sequencing, how it differs from Sanger sequencing, and the different steps involved in sequencing a DNA sample with NGS methods, including library preparation, Sequencing by Synthesis and the sequencing reaction, cluster generation from fragments, indexing, filtering, mapping, and demultiplexing the reads.

### *How to sequence the human genome*

**Title**: How to sequence the human genome

**Presenter(s)**: Mark J. Kiel

**Length**: 5:04

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=MvuYATh7Y74")
```

Link to video transcript [here](chapter3.2_transcript.qmd#sec-video1).

### *Next Generation Sequencing: A Step-by-Step Guide to DNA Sequencing*

**Title**: Next Generation Sequencing: A Step-by-Step Guide to DNA Sequencing

**Description**: Next Generation Sequencing (NGS) is used to sequence both DNA and RNA. Billions of DNA strands get sequenced simultaneously using NGS. Whereas with Sanger Sequencing, only one strand is sequenced at a time. While the Human Genome Project took over 30 years to sequence the human genome for the first time. Now with Next Generation Sequencing, a whole human genome can be sequenced in just one day. This video describes the library preparation process, cluster generation, the sequencing reaction and filtering and alignment of the sequencing data. This video focuses on the primary sequencing method used by Illumina, Sequencing by Synthesis (SBS).

**Presenter(s)**: ClevaLab

**Length**: 7:37

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=WKAUtJQ69n8")
```

Link to video transcript [here](chapter3.2_transcript.qmd#sec-video2).
