---
title: "Chapter 8: Post-GWAS bioinformatics"
---

------------------------------------------------------------------------

# 8.1 SNP Heritability {#sec-section1}

### 

> Goals of this section:
>
> 1.  

#### 

**Title**: Heritability and SNP Heritability

**Description**:

**Presenter(s)**: Alkes Price, Broad Institute

**Level**:

**Length**: 56:27

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=VDqbWa028io&t=598s")
```

Link to video transcript [here](chapter8.1_transcript.qmd).

------------------------------------------------------------------------

# 8.2 Genetic Correlations and Partitioned LDSC {#sec-section2}

> Goals of this section:
>
> 1.  

**Title**: Genetic Correlation and Partitioning

**Description**:

**Presenter(s)**: Patrick Turley, ATGU, Massachusetts General Hospital

**Level**: Beginner friendly

**Length**: 12:11

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=yHPfoIKYK6w&t=295s")
```

Link to video transcript [here](chapter8.2_transcript.qmd).

------------------------------------------------------------------------

# 8.3 Gene association analysis {#sec-section3}

> Goals of this section:
>
> 1.  

## MAGMA

**Title**: How do we go from genetic discoveries from GWAS/WGS/WES to mechanistic disease insight?

**Description**:

**Presenter(s)**: Danielle Posthuma

**Level**: Beginner friendly

**Length**: 9:53

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=RQwASJcR8_4")
```

Link to video transcript [here](chapter8.3_transcript.qmd#sec-video1).

------------------------------------------------------------------------

## FUMA

**Title**: FUMA: Functional mapping and annotation of genetic associations

**Description**:

**Presenter(s)**: Kyoko Watanabe

**Level**: Beginner friendly

**Length**: 14:19

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=fLEooAu3VwM")
```

Link to video transcript [here](chapter8.3_transcript.qmd#sec-video2).

------------------------------------------------------------------------

# 8.4 Gene set analysis {#sec-section4}

> Goals of this section:
>
> 1.  

**Title**: How do we go from genetic discoveries from GWAS/WGS/WES to mechanistic disease insight?

**Description**:

**Presenter(s)**: Danielle Posthuma

**Level**: Beginner friendly

**Length**: 9:53

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=RQwASJcR8_4")
```

Link to video transcript [here](chapter8.3_transcript.qmd#sec-video1).

[\
](https://www.youtube.com/watch?v=oxFqpv-4cxo)

------------------------------------------------------------------------

# 8.5 Fine-mapping {#sec-section5}

> Goals of this section:
>
> 1.  

**Title**: Introduction to fine-mapping methods

**Description**:

**Presenter(s)**: Hilary Finucane

**Level**: Beginner friendly

**Length**: 52:12

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=S6vfOr336b0")
```

Link to video transcript [here](chapter8.5_transcript.qmd).

------------------------------------------------------------------------

# 8.6 Quantitative Trait Loci (QTLs) {#sec-section6}

### General Introduction to GWAS Studies and eQTL Analysis

> Goals of this section:
>
> 1.  Be able to define GWAS and understand the general genetic architecture of common risk variants
>
> 2.  Be able to define an eQTL (Expression Quantitative Trait Loci) and how they may be useful to interprating GWAS results

#### **GWAS Studies and eQTL Analysis**

**Title**: GWAS Studies and eQTL Analysis.

**Description**: This short video briefly reviews GWAS studies and the GWAS catalog resource in order to discuss a main caveat of GWAS - that most significant trait associated variants lay in non-coding regions of the genome with putative regulatory function. The video then presents expression quantitative trait loci (eQTLs) as a method to connect SNPs to gene regulatory effects in a tissue-specific manner.

**Presenter(s)**: Xiaole Shirley Liu

**Level**: Beginner

**Length**: 14:23

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=ARdlw7irNOw&ab_channel=XiaoleShirleyLiu")
```

Link to video transcript [here](chapter8.6_transcript.qmd#sec-video1).

------------------------------------------------------------------------

### Expression Quantitative Trait loci (eQTLs)

> Goals of this section:
>
> 1.  Obtain a solid understanding of eQTLs and related analytical approaches.
>
> 2.  Examine the role of eQTLs in the relationship of genes and molecular networks with phenotypic QTLs (pQTLs).
>
> 3.  eQTL for co-expression networks.

#### Introduction to expression quantitative trait loci

**TItle**: MPG Primer: Introduction to expression quantitative trait loci (2021).

**Description**: This video introduces expression quantitative trait loci (eQTL) analysis, discusses how to address batch effects and covariate correction, and control for false discovery rate.

**Presenter(s)**: Francois Auget - Broad Institute

**Level**: intermediate (the content will be easier to follow by those who already went through some introductory materials on GWAS and gene expression)

**Length**: 52:43

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=24U8lyqu65A")
```

Link to video transcript [here](chapter8.6_transcript.qmd#sec-video2).

------------------------------------------------------------------------

**Title**: Introduction to expression (e)QTL & their role in connecting QTL to genes and molecular networks.

**Description**: This video explains the concepts of transcriptome-wide association studies, gene expression prediction related to phenotypes of interest, and accompanying statistical tests.

**Level**: intermediate (the content will be easier to follow by those who already went through some introductory materials on GWAS and gene expression)

**Author**: Laura Saba

**Length**: 1:26:43

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=8jiNHuOgr1A&ab_channel=opar")
```

Link to video transcript [here](chapter8.6_transcript.qmd#sec-video3).

------------------------------------------------------------------------

### From SNPs to Genes

**Video 4**: Linking SNPs with genes in GWAS. This video discuss the relevance of gene regulation in GWAS, explains transcriptome-wide association studies and co-localization of GWAS with eQTLs as methods to link SNPs to gene expression and trait associations, discusses proximity and the abstrat mediation model, and covered combining multiple SNP-to-gene strategies based on the Nature Genetics Publication: "Combining SNP-to-gene linking strategies to identigy diease genes and assess disease omonigenicity".

**Level**: intermediate (the content will be easier to follow by those who already went through some introductory materials on GWAS and gene expression)

**Author**: Luke O'Connor and Dan Weiner, Broad Institute

**Length**: 47:52

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=2O89Cl8RCe0")
```

Link to video transcript [here](chapter8.6_transcript.qmd#sec-video4).

------------------------------------------------------------------------

### Resources

<https://chr1swallace.github.io/coloc/articles/a01_intro.html> <http://gusevlab.org/projects/fusion/>

------------------------------------------------------------------------

# 8.7 TWAS/PWAS/MWAS {#sec-section7}

### Transcriptome-wide association studies (TWAS)

> Goals of this section:
>
> 1.  Obtain

#### Understanding GWAS mechanisms with Transcriptome-Wide Association Studies

**Title**: Understanding GWAS mechanisms with Transcriptome-Wide Association Studies

**Description**: This video explains the concepts of transcriptome-wide association studies, gene expression prediction related to phenotypes of interest, and accompanying statistical tests.

**Presenter(s)**: Sasha Gusev (alexander_gusev\@dfci.harvard.edu)

**Level**: Intermediate (the content will be easier to follow by those who already went through some introductory materials on GWAS and gene expression)

**Length**: 41:01

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=cfEGf6ezR-c")
```

Link to video transcript [here](chapter8.7_transcript.qmd#sec-video1).

------------------------------------------------------------------------

#### Methods for GWAS/TWAS

**Title**:

**Description**: A very brief introduction to what TWAS studies are.

**Presenter(s)**: John Eicher (https://www.researchgate.net/profile/John-Eicher-2) and Noah Zaitlen (nzaitlen\@ucla.edu)

**Level**: Beginner

**Length**: 2:40

Link to video transcript [here](chapter8.7_transcript.qmd#sec-video2).

------------------------------------------------------------------------

#### TWAS Primer

**Title**: PGC TWAS Primer

**Description**: This video provides a brief background to TWAS and how to run it using the FUSION software.

**Presenter(s)**: Sasha Gusev (alexander_gusev\@dfci.harvard.edu)

**Level**: Beginner to Intermediate

**Length**: 16:39

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=BhZ0NUGwZJg")
```

Link to video transcript [here](chapter8.7_transcript.qmd#sec-video3).

------------------------------------------------------------------------

### MWAS tutorial

> Goals of this section:
>
> 1.  Obtain

**Title**:

**Description**: It is a private video hosted on the Edinburgh University cloud so I cannot access it - to be updated when the owner grants access

**Presenter(s)**: Xueyi Shen (xueyi.shen\@ed.ac.uk)

**Level**:

Link to video transcript [here](chapter8.7_transcript.qmd#sec-video4).

------------------------------------------------------------------------

# 8.8 PheWAS {#sec-section8}

> Goals of this section:
>
> 1.  

**Title**: PheWAS: Discovering gene-disease associations

**Description**: Overview of PheWAS. This video talks about how PheWAS and GWAS could work in concert with each other and what data are used for the phenome. After 9:45, Dr. Josh Denny pivots the focus from PheWAS to phenotypic risk score (PheRS).

**Presenter(s)**: Dr. Joshua Denny, All of Us Research Program

**Level**: Intermediate

**Length**: 13:40

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=8sSJPHZGmYY")
```

Link to video transcript [here](chapter8.8_transcript.qmd#sec-video1).

------------------------------------------------------------------------

**Title**: Using EHR-based genomic approaches to understand the relationship between mental and physical health

**Description**: Example of PheWAS and how to interpret the PheWAS results (starting 28:15). **Note**: There is no current video on *LabWAS* yet, but the concept is similar to PheWAS and in this youtube video, Dr. Lea Davis talked briefly on the QC of the lab results in the BioVU Biobank (starting 39:00).

**Presenter(s)**: Dr. Lea Davis, Vanderbilt University

**Level**: Intermediate

**Length**: 58:40

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=JirCyMOC8fI")
```

Link to video transcript [here](chapter8.8_transcript.qmd#sec-video2).
