---
title: "Chapter 4: Study Designs"
---

> *Chapter goals:*
>
> 1.  *Understand the different types of epidemiological study designs and when their use is appropriate.*
>
> 2.  *Understand different sources of error that can occur in epidemiological studies and how to adjust/correct for that error.*
>
> 3.  *Gain insight into different types of genetic study designs and how to select study designs that are appropriate to your research goals.*

## 4.1 Epidemiological Study Design {#sec-section1}

> This video by Drs. Appuhamy and Clark gives an overview of different epidemiological study designs, covering the basics of the study types: ecological study, case series, cross-sectional study, case-control study, cohort study, interventional study, systematic review, and meta-analysis.

**Title**: Epidemiological Studies: A Beginners guide

**Presenter(s)**: Ranil Appuhamy, MD (World Health Organization)

**Length**: 9:42

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=Jd3gFT0-C4s")
```

Link to video transcript [here](chapter4.1_transcript.qmd).

------------------------------------------------------------------------

## 4.2 Confounding, Chance, and Bias {#sec-section2}

> This section includes a video from Cochrane Austria that provides a fairly comprehensive overview of potential sources of error for clinical and genetic studies. The video discusses "what is confounding?", different types of confounders, and how to deal with confounding, chance and random error, and different types of biases: selection bias, performance bias, measurement bias, and attrition bias. 

**Title**: Confounding, Chance, and Bias

**Presenter(s)**: Cochrane Austria (Department for Evidence-based Medicine and Evaluation, Danube University Krems)

**Length**: 9:50

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=bcfg9kcxeuU")
```

Link to video transcript [here](chapter4.2_transcript.qmd).

------------------------------------------------------------------------

## 4.3 Genetic Study Designs {#sec-section3}

> This section explores different types of genetic study designs, starting with a video from OpenLearn on twin studies and how studies of twins allow for detection and distinguishing of genetic- versus environmentally- affected traits.

**Title**: What are "Twin studies"?

**Presenter(s)**: OpenLearn from the Open University

**Length**: 2:13

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=BTYCv1ObZrI&t=1s")
```

Link to video transcript [here](chapter4.3_transcript.qmd#sec-video1).
