---
title: "Chapter 2: The Genome"
---

------------------------------------------------------------------------

## 2.1 Organization of the genome {#sec-section1}

**Title**: An Introduction to the Human Genome

**Description**: A brief overview of the human genome: What is DNA? How is DNA organized? How is DNA regulated? What are genes?

**Presenter(s)**: HMX Genetics, Harvard University

**Length**: 5:35

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=jEJp7B6u_dY")
```

Link to video transcript [here](chapter2.1_transcript.qmd).

------------------------------------------------------------------------

## 2.2 Genetic variation {#sec-section2}

**Title**: Genetic Variation and Mutation

**Description**: A brief overview on genetic variation (SNPs, CNVs)

**Presenter(s)**: Precision Health

**Length**: 4:59

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=7W17xH62MVA")
```

Link to video transcript [here](chapter2.2_transcript.qmd).

------------------------------------------------------------------------

## 2.3 Evolutionary signatures {#sec-section3}

### Origins of Genetic Variation

**Title**: Origins of Genetic Variation

**Description**: A brief overview on genetic variation (SNPs, CNVs)

**Presenter(s)**: Jessica Pamment, Professor, DePaul University

**Length**: 4:47

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=mblvxhkXjXA")
```

Link to video transcript [here](chapter2.3_transcript.qmd#sec-origins).

------------------------------------------------------------------------

### Natural Selection & Human Genetic Variation

**Title**: MPG Primer: Natural selection & human genetic variation

**Description**: A brief overview on genetic variation (SNPs, CNVs)

**Presenter(s)**: Stephen Schaffner, Computational Biologist, Broad Institute

**Length**: 54:09

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=QerLgeslHdI")
```

Link to video transcript [here](chapter2.3_transcript.qmd#sec-selection).

------------------------------------------------------------------------

## 2.4 Linkage disequilibrium {#sec-section4}

### What is linkage disequilibrium?

**Title**: What is linkage disequilibrium?

**Description**:

**Presenter(s)**: Gábor Mészáros, PhD

**Length**: 12:52

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=C3MYoasLSHQ")
```

Link to video transcript [here](chapter2.4_transcript.qmd#sec-video1).

------------------------------------------------------------------------

### Measuring linkage disequilibrium

**Title**: How to measure linkage disequilibrium?

**Description**:

**Presenter(s)**: Gábor Mészáros, PhD

**Length**: 11:31

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=vJZUnflN7kk")
```

Link to video transcript [here](chapter2.4_transcript.qmd#sec-video2).

------------------------------------------------------------------------

### (Advanced) Measuring linkage disequilibrium

**Title**: How to measure linkage disequilibrium? (ADVANCED)

**Description**:

**Presenter(s)**: Gábor Mészáros, PhD

**Length**: 9:28

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=65R1qUmXleg")
```

Link to video transcript [here](chapter2.4_transcript.qmd#sec-video3).

------------------------------------------------------------------------

### Computing linkage disequilibrium

**Title**: Compute linkage disequilibrium (Part 1)

**Description**:

**Presenter(s)**: Gábor Mészáros, PhD

**Length**: 10:18

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=3cNmTwm-x-Q&t=6s")
```

Link to video transcript [here](chapter2.4_transcript.qmd#sec-video4).
