---
title: "Chapter 10: Other Considerations"
---

> *Chapter goals:*
>
> 1.  *Gain insight into different career paths in psychiatric genetics*
>
> 2.  *Understand the limitations of genetic prediction*
>
> 3.  *Understand how small effect sizes in genetic studies are still beneficial for understanding disease etiology.*
>
> 4.  *Understand what biological essentialism is, how it impacts medicine and science, exacerbates health disparities and perpetuates discrimination.*
>
> 5.  *Understand the importance of diversity in genetic studies, how to include and protect vulnerable populations in genetic studies, and how to conduct equitable collaborative research with low- or middle-income countries (LMIC).*
>
> 6.  *Gain insight into the structure and rules of the European General Data Protection Regulation (GDPR), why it is important, and how to navigate the guidelines for scientific research.*

# 10.1 A Career in Psychiatric Genetics {#sec-section1}

> In this video, Drs. Lewis and Curmi discuss different genetics topics, careers in psychiatric genetics, and the future of psychiatric genetics, from both data/research and clinical viewpoints.

**Title**: How does Genetics Affect our Mental Health?

**Presenter(s)**:

-   Cathryn Lewis, PhD (Social, Genetic, and Developmental Psychiatry Centre, King's College London)

-   Alex Curmi, MD (Maudsley NHS Foundation Trust)

**Level**: Beginner

**Length**: 58:54

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=8jtoLRmC5F4")
```

Link to video transcript [here](chapter10.1_transcript.qmd).

------------------------------------------------------------------------

# 10.2 Caution in Genetic Prediction {#sec-section2}

> In this brief video, Dr. Edenberg discusses the limitations of genetic prediction, specifically polygenic risk scores, in the field of psychiatry.

**Title**: Predicting the likelihood of future psychiatric disorders: a closer look, and some cautions.

**Presenter(s)**: Howard Edenberg, PhD (Department of Biochemistry and Molecular Biology, Indiana University School of Medicine)

**Level**: Beginner

**Length**: 4:30

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=P8Odk5HuY50")
```

Link to video transcript [here](chapter10.2_transcript.qmd).

------------------------------------------------------------------------

# 10.3 Small Effect Sizes {#sec-section3}

> In this video, Dr. Edenberg briefly discusses the benefits of studying genetic variants with small effect sizes in genome-wide association studies, and how understanding how common variants contribute to disease risk, despite having very small effect sizes, still leads to better understanding of disease.

**Title**: Small effect sizes

**Presenter(s)**: Howard Edenberg, PhD (Department of Biochemistry and Molecular Biology, Indiana University School of Medicine)

**Level**: Beginner

**Length**: 4:05

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=l5tsUyhXjt4")
```

Link to video transcript [here](chapter10.3_transcript.qmd).

------------------------------------------------------------------------

# 10.4 Language in Genetics Research {#sec-section4}

***Video Content coming soon!***

**Presenter(s)**:

-   Carina Seah, MS (Department of Genetics and Genomic Sciences, Icahn School of Medicine at Mount Sinai)

-   Kayla Townsley, MS (Department of Genetics and Genomic Sciences, Icahn School of Medicine at Mount Sinai)

**Level**: Beginner

------------------------------------------------------------------------

# 10.5 Equitable Collaboration for LMIC Researchers {#sec-section5}

***Video Content coming soon!***

Please check out these papers in the meantime:

-   Fatumo, S., Chikowore, T., Choudhury, A. *et al.* (2022): A roadmap to increase diversity in genomic studies. *Nat Med* **28**, 243--250 <https://doi.org/10.1038/s41591-021-01672-4>

-   Palk A, Illes J, Thompson PM, Stein DJ (2020): Ethical issues in global neuroimaging genetics collaborations. NeuroImage 221: 117208: <https://doi.org/10.1016/j.neuroimage.2020.117208>

-   S.R. Carroll, I. Garba, R. Plevel, D. Small-Rodriguez, V.Y. Hiratsuka, M. Hudson, N.A. Garrison (2022): Using Indigenous Standards to Implement the CARE Principles: Setting Expectations through Tribal Research Codes. Front Genet, 13 Article 823309: <https://doi.org/10.3389/fgene.2022.823309>

------------------------------------------------------------------------

# 10.6 GDPR for Dummies: A Survival Guide for Genetics Research {#sec-section6}

> In this video, Heidi Bentzen gives a comprehensive overview of the European General Data Protection Regulation (GDPR): the purpose of the GDPR, and how to ensure that your research is compliant with GDPR statutes.

**Title**: GDPR for dummies: A survival guide for genetics research

**Presenter(s)**: Heidi Beate Bentzen, LLM (Centre for Medical Ethics, University of Oslo)

**Level**: Beginner

**Length**: 23:46

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=bxCV2Z5i_yI&t=8s")
```

Link to video transcript [here](chapter10.6_transcript.qmd).

Link to the [GDPR text](https://gdpr-info.eu/).
