---
title: "Chapter 1: Introduction"
---

> *Chapter goals:*
>
> 1.  *Gain a basic understanding of psychiatric genetics and genome-wide association studies.*
> 2.  *Understand the recent history of psychiatric genetics.*
> 3.  *Gain insight into the current state of psychiatric disorder research.*

## 1.1 What are Psychiatric Disorders? {#sec-section1}

> The video below from the Broad Institute gives a very brief overview of Psychiatric disorder genetics.

**Title**: Psychiatric Disorders

**Presenter(s)**: Broad Institute of MIT and Harvard

**Level**: Beginner

**Length**: 2:42

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=7j29y8uQc5E")
```

Link to video transcript [here](chapter1.1_transcript.qmd).

------------------------------------------------------------------------

## 1.2 Epidemiology {#sec-section2}

> Dr. Andlauer gives a brief overview of genome-wide association studies and genetic epidemiology.

**Title**: What is a genome-wide association study?

**Presenter(s)**: Till Andlauer, PhD (Department of Computational Biology and Digital Sciences, Boehringer Ingelheim)

**Level**: Beginner

**Length**: 3:15

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=LxUXinawAv0")
```

Link to video transcript [here](chapter1.2_transcript.qmd).

------------------------------------------------------------------------

## 1.3 History {#sec-section3}

> In the videos below, Dr. Kendler gives brief overview of some of the prominent players in psychiatry and psychiatric genetics, starting from the late 1700s through the mid-1900s. Part 1 highlights the work of Robert Burton, Prosper Lucas, Jenny Koller, Ernst Rüdin, and Eugen Bleuler, and discusses the darker periods of the history of psychiatric genetics during the late 1920's and 1930's.
>
> Part 2 is a deeper dive into a single episode in the history of Psychiatry: "The Beginnings of the Debate Between the Mendelians and the Biometricians in Psychiatric Genetics", focusing on four individuals: David Heron, Karl Pearson, Aaron Rosanoff, and Charles Davenport during 1913 to 1914.

**Title**: History of Psychiatric Genetics (Part 1)

**Presenter(s)**: Ken Kendler, MD (Virginia Institute for Psychiatric and Behavioral Genetics, Virginia Commonwealth University)

**Length**: 21:44

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=ikJ7zyo4Spk")
```

Link to video transcript [here](chapter1.3_transcript.qmd#sec-part1).

------------------------------------------------------------------------

**Title**: History of Psychiatric Genetics (Part 2)

**Presenter(s)**: Ken Kendler, MD (Virginia Institute for Psychiatric and Behavioral Genetics, Virginia Commonwealth University)

**Length**: 43:41

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=Xwhd-rTPXvg")
```

Link to video transcript [here](chapter1.3_transcript.qmd#sec-part2).

------------------------------------------------------------------------

## 1.4 Psychiatric Genomics: State-of-the-Science {#sec-section4}

> Below are videos describing the current state of psychiatric research. These are state-of-the-science videos for psychiatric disorders, which were recorded for the PGC by early career researchers, with accompanying papers in [*Psychological Medicine* (Vol 51, issue 13)](https://www.cambridge.org/core/journals/psychological-medicine/issue/F270E99C61D43EABAA4DFEF0CA154B06).

### The Psychiatric Genomics Consortium

**Title**: PGC: Future Perspectives

**Presenter(s)**: Patrick Sullivan, MD, FRANZCP (Department of Genetics, Department of Psychiatry, University of North Carolina at Chapel Hill)

**Length**: 33:33

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=ZhqMgazsRXM")
```

Link to video transcript [here](chapter1.4_transcript.qmd#sec-psychgen).

------------------------------------------------------------------------

### *ADHD*

**Title**: Insights into Attention-Deficit/Hyperactivity Disorder from Genetic Studies

**Presenter(s)**:

-   Joanna Martin, PhD (Division of Psychological Medicine and Clinical Neurosciences, Cardiff University School of Medicine)

**Coauthor(s)**:

-   Christie Burton, PhD (Department of Psychiatry, Hospital for Sick Children, University of Toronto)

-   Isabell Brikell, PhD (Department of Medical Epidemiology and Biostatistics, Karolinska Institutet)

-   Nina Roth Mota, PhD (Department of Human Genetics, Radboud University Medical Center, Donders Institute for Brain, Cognition and Behaviour)

**Length**: 14:24

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=I1X6Ce23buk")
```

Link to video transcript [here](chapter1.4_transcript.qmd#sec-adhd).

------------------------------------------------------------------------

### *Alzheimer's Disease*

See [Dementia](chapter1.qmd#sec-dementia)

------------------------------------------------------------------------

### *Anxiety Disorders*

**Title**: Genetic contributions to anxiety disorders: Where we are and where we are heading

**Presenter(s)**:

-   Kirstin Purves, PhD (Our Future Health)

-   Daniel Levey, PhD (Department of Psychiatry, Yale University School of Medicine)

-   Rosa Cheesman, PhD (Department of Psychology, University of Oslo)

**Coauthor(s)**:

-   Helga Ask, PhD (Department of Mental Disorders, Norwegian Institute of Public Health)

-   Heike Weber, PhD (Department of Psychiatry, Psychosomatic Medicine and Psychotherapy, Goethe-Universität Frankfurt am Main)

-   Eshim S. Jami, PhD (Department of Clinical Education and Health Psychology, University College London)

**Length**: 11:32

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=4Cn2TFSzIz4")
```

Link to video transcript [here](chapter1.4_transcript.qmd#sec-anxiety).

------------------------------------------------------------------------

### *Autism Spectrum Disorder*

**Title**: Genetic contributions to Autism Spectrum Disorder

**Presenter(s)**:

-   Alexandra Havdahl, PhD (Department of Psychology, University of Oslo)

-   Anna Starnawska, PhD (Department of Biomedicine, Aarhus University)

-   Maria Niarchou, PhD (Department of Medicine, Vanderbilt University)

**Coauthor(s)**:

-   Mohammed Uddin, PhD (Mohammed Bin Rashid University of Medicine and Health Sciences, College of Medicine, Dubai)

-   Celia van der Merwe, PhD (Analytic and Translatoinal Genetics Unit, Broad Institute of MIT and Harvard)

-   Varun Warrier, PhD (Autism Research Centre, University of Cambridge)

**Length**: 16:42

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=TFyjn-gFCt8")
```

Link to video transcript [here](chapter1.4_transcript.qmd#sec-asd).

------------------------------------------------------------------------

### *Bipolar Disorder*

**Title**: Genetics of Bipolar Disorder

**Presenter(s)**: Brandon Coombes, PhD (Division of Computational Biology, Department of Quantitative Health Sciences, Mayo Clinic)

**Coauthor(s)**: Kevin O'Connell, PhD (NORMENT Centre, Institute of Clinical Medicine, University of Oslo)

**Length**: 16:35

```{r, echo=FALSE}
library("vembedr")
embed_url("https://youtu.be/Z4e0SagGncY")
```

Link to video transcript [here](chapter1.4_transcript.qmd#sec-bipolar).

Paper referenced in the talk:

> O'Connell KS, Coombes BJ. **Genetic contributions to bipolar disorder: current status and future directions**. *Psychol Med*. 2021 Oct;51(13):2156-2167. doi: [10.1017/S0033291721001252](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8477227/).

------------------------------------------------------------------------

### *Dementia* {#sec-dementia}

**Title**: Genetic Risk for Dementia

**Presenter(s)**: Malia Rumbaugh, MS (Department of Medical and Molecular Genetics, Indiana University-Purdue University Indianapolis)

**Length**: 58:02 (Frontotemporal dementia starts at 17:55)

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=PBiDpFaV5_g&t=1337s")
```

Link to video transcript [here](chapter1.4_transcript.qmd#sec-dementia).

------------------------------------------------------------------------

### *Eating Disorders*

**Title**: Genetics of Eating Disorders

**Presenter(s)**:

-   Helena Davies, MSc (Social, Genetic, and Developmental Psychiatry Centre, King's College London)

-   Alish Palmos, PhD (Social, Genetic, and Developmental Psychiatry Centre, King's College London)

-   Zeynep Yilmaz, PhD (National Centre for Register-based Research, Aarhus University)

**Coauthor(s)**:

-   Hunna Watson, PhD (Department of Psychiatry, University of North Carolina at Chapel Hill)

-   Jessica Baker, PhD (Center of Excellence for Eating Disorders, University of North Carolina at Chapel Hill)

-   Avina Hunjan, PhD (Social, Genetic, and Developmental Psychiatry Centre, King's College London)

**Length**: 21:30

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=VyNaE02UW20")
```

Link to video transcript [here](chapter1.4_transcript.qmd#sec-eatingdisorders).

------------------------------------------------------------------------

### *Major Depression*

**Title**: Major Depressive Disorder: Introduction and General Epidemiology

**Presenter(s)**:

-   Kimberley Kendall, MD, PhD (Division of Psychological Medicine and Clinical Neurosciences, Cardiff University)

-   Lu Yi, PhD (Department of Medical Epidemiology and Biostatistics, Karolinska Institutet)

-   Eva Schulte, MD, PhD (Institute of Psychiatric Phenomics and Genomics, University of Munich)

-   Till Andlauer, PhD (Department of Computational Biology and Digital Sciences, Boehringer Ingelheim)

-   Jurgen Luykx, MD, PhD (Human Neurogenetics Unit, University Medical Center Utrecht)

-   Karmel Choi, PhD (Department of Psychiatry, Massachusetts General Hospital)

-   Evelien Van Assche, MD, PhD (Department of Psychiatry, University of Münster)

**Length**: 32:13

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=-NQ5-MJaea0")
```

Link to video transcript [here](chapter1.4_transcript.qmd#sec-mdd).

------------------------------------------------------------------------

### *Obsessive Compulsive Disorder*

**Title**: Genetics of Obsessive-Compulsive Disorder: What we know in 2020

**Presenter(s)**: Christie Burton, PhD (Department of Psychiatry, Hospital for Sick Children, University of Toronto)

**Coauthor(s)**:

-   Behrang Mahjani, PhD (Department of Psychiatry, Icahn School of Medicine at Mount Sinai)

-   Katharina Bey, PhD (Clinic and Polyclinic for Psychiatry and Psychotherapy, University Hospital Bonn)

-   Julia Boberg, PhD (Department of Clinical Neuroscience, Karolinska Institutet)

**Length**: 15:28

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=_VVdqHOr7v0")
```

Link to video transcript [here](chapter1.4_transcript.qmd#sec-ocd).

------------------------------------------------------------------------

### *Post-Traumatic Stress Disorder*

**Title**: Posttraumatic Stress Disorder: From Gene Discovery to Disease Biology

**Presenter(s)**: Frank Wendt, PhD (Department of Anthropology, University of Toronto)

**Length**: 15:33

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=jy2StAoWjzs")
```

Link to video transcript [here](chapter1.4_transcript.qmd#sec-ptsd).

------------------------------------------------------------------------

### *Schizophrenia*

**Title**: Genetic Architecture of Schizophrenia

**Presenter(s)**:

-   Kaarina Kowalec, PhD (College of Pharmacy, University of Manitoba)

-   Adeniran Okewole, MBBS, FWACP, FMCPsy (Department of Clinical Services, Neuropsychiatric Hospital Aro)

-   Sathish Periyasamy, PhD (Queensland Centre for Mental Health Research, The University of Queensland, University of Queensland)

-   Marcos Santoro, PhD (Departamento de Bioquímica, Universidade Federal de São Paulo, Brazil)

-   Sophie Legge, PhD (Division of Psychological Medicine and Clinical Neurosciences, Cardiff University)

**Coauthor(s)**: Arsalan Arsalan, MD (Department of Pharmacy, University of Peshawar, Pakistan)

**Length**: 29:52

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=NMr5sLt3U4o")
```

Link to video transcript [here](chapter1.4_transcript.qmd#sec-schizophrenia).

------------------------------------------------------------------------

### *Shared Genetic Architecture*

**Title**: Shared Genetic Architecture across Psychiatric Disorders

**Presenter(s)**: Andrew Grotzinger, PhD (Institute for Behavioral Genetics, University of Colorado Boulder)

**Length**: 14:04

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=XwBoFzdz3ts")
```

Link to video transcript [here](chapter1.4_transcript.qmd#sec-sharedgenetic).

------------------------------------------------------------------------

### *Substance Use Disorder*

**Title**: The Genetics of Substance Use Disorders: A Review

**Presenter(s)**:

-   Joseph Deak, PhD (Department of Psychiatry, Yale University School of Medicine)

-   Emma Johnson, PhD (Department of Psychiatry, Washington University in St. Louis School of Medicine)

**Length**: 17:38

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=NdOR9NxMH2Q")
```

Link to video transcript [here](chapter1.4_transcript.qmd#sec-sud).

Paper referenced in the talk:

> Deak JD, Johnson EC. **Genetics of substance use disorders: a review**. *Psychol Med*. 2021 Oct;51(13):2189-2200. doi: [10.1017/S0033291721000969](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8477224/)

------------------------------------------------------------------------

### *Suicidal Thoughts and Behaviours*

**Title**: Insight into the Genetics of Suicide

**Presenter(s)**:

-   Hilary Coon, PhD (Department of Psychiatry, University of Utah)

-   Anna Docherty, PhD (Department of Psychiatry, University of Utah)

-   Douglas Ruderfer, PhD (Division of Genetic Medicine, Vanderbilt University)

-   Niamh Mullins, PhD (Department of Psychiatry, Icahn School of Medicine at Mount Sinai)

**Length**: 58:54

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=uQAO_XwN9Fk&t=80s")
```

Link to video transcript [here](chapter1.4_transcript.qmd#sec-suicide).

------------------------------------------------------------------------

### *Tourette Syndrome*

**Title**: Genetics of Tourette Syndrome

**Presenter(s)**: Matt Halvorsen, PhD (Department of Genetics, University of North Carolina School of Medicine)

**Length**: 15:23

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=SpRczJwXkpc")
```

Link to video transcript [here](chapter1.4_transcript.qmd#sec-tourette).
