---
title: "Chapter 7: Ancestry-Specific Analyses and Considerations"
---

------------------------------------------------------------------------

# Chapter 7.1: Cross-ancestry analysis {#sec-section1}

**Title**: Cross-ancestry PTSD and Polygenic Findings, and the Cross-Population SIG of the PGC

**Description**:

**Presenter(s)**: Laramie Duncan

**Length**: Video starts at 00:00 and goes until 13:39 (13:39 total)

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=p3m6MM5b-UU&t=148s")
```

Link to video transcript [here](chapter7_transcript.qmd#sec-video1).

------------------------------------------------------------------------

# Chapter 7.2: Ancestry-specific PRS {#sec-section2}

**Title**: Clinical use of current polygenic risk scores may exacerbate health disparities

**Description**:

**Presenter(s)**: Alicia Martin

**Length**: Video starts at 32:39 and goes to 47:06 (15:45 total)

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=p3m6MM5b-UU&t=2321s")
```

Link to video transcript [here](chapter7_transcript.qmd#sec-video2).

------------------------------------------------------------------------

# Chapter 7.3: Local Ancestry and Admixed Populations {#sec-section3}

**Title**: Tractor: A framework enabling the well-calibrated genomic analysis of psychiatric traits across admixed populations

**Description**:

**Presenter(s)**: Elizabeth Atkinson

**Length**: Video starts at 14:50 and ends at 32:38 (total 19:28)

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=p3m6MM5b-UU&t=890s")
```

Link to video transcript [here](chapter7_transcript.qmd#sec-video3).
