---
title: "Chapter 6: Polygenic Scores"
---

------------------------------------------------------------------------

# 6.1 Polygenic Risk Scores

Goals of this section:

> 1.  Understand "what is a polygenic risk score?"
>
> 2.  Understand the basic principals of calculating polygenic risk scores
>
> 3.  Understand different polygenic risk scoring methods, their benefits, and pitfalls.

## Polygenic Risk Scores: Introduction

**Title**: Polygenic Risk Scores

**Description**: Polygenic risk scores in a nutshell. This video is designed to explain PRS to a lay audience and answers the question why we care about PRS and what we could do with the knowledge of PRS to improve human health.

**Presenter(s)**: Center for Personalized Medicine, Oxford University

**Level**: Beginner friendly

**Length**: 4:03

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=1I64mnEcIVA")
```

Link to video transcript [here](chapter6.1_transcript.qmd#sec-video1).

------------------------------------------------------------------------

**Title**: Polygenic Risk Scores

**Description**: Introduction to polygenic risk scores. This video explains how to calculate PRS and how to interpret PRS in a clear and easily understandable way.

**Presenter(s)**: Till Andlauer

**Level**: Beginner friendly

**Length**: 4:36

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=WlcEADfD-HU")
```

Link to video transcript [here](chapter6.1_transcript.qmd#sec-video2).

------------------------------------------------------------------------

## Polygenic Risk Scores: In detail

**Title**: Polygenic risk scores

**Description**: A more detailed introduction to polygenic risk scores

**Presenter(s)**: Adrian Campos, Adrian.Campos\@qimrberghofer.edu.au

**Level**: Intermediate

**Length**: 22:21

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=FKw2XltZSpY&t=1336s")
```

Link to video transcript [here](chapter6.1_transcript.qmd#sec-video3).

------------------------------------------------------------------------

## Polygenic Scoring Methods

**Title**: Polygenic Scoring Methods: Comparison and Implementation

**Description**: Comparison of different PRS methods I. This video compares different available PRS methods on their performance and required computational time (ends at 12:30) in the UK Biobank. Towards the end, Dr. Pain introduces their new software to calculate standardized individual polygenic risk score based on genetic ancestry.

**Presenter(s)**: Oliver Pain

**Level**: Intermediate

**Length**: 21:14

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=jTlz5z-alts")
```

Link to video transcript [here](chapter6.1_transcript.qmd#sec-video4).

------------------------------------------------------------------------

**Title**: Polygenic risk scores: PGS comparison

**Description**: This is another video comparing different PRS methods on their performance in the cohorts in the Psychiatric Genomics Consortium (i.e., schizophrenia cohorts). Dr. Ni gives toy examples to explain the different parameters they used to evaluate how well each PRS method performed.

**Presenter(s)**: Guiyan Ni

**Level**: Intermediate

**Length**: 13:52

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=H18i2Oe05Gw")
```

Link to video transcript [here](chapter6.1_transcript.qmd#sec-video5).
