---
title: "Chapter 4: Study designs"
---

------------------------------------------------------------------------

## 4.1 Epidemiological study design {#sec-section1}

**Title**: Epidemiological Studies: A Beginners guide

**Description**: This video gives a simple overview of the most common types of epidemiological studies, their advantages and disadvantages. These include ecological, case-series, case control, cohort and interventional studies. It also looks at systematic reviews and meta-analysis.

**Presenter(s)**: Ranil Appuhamy, James Clark, Let's Learn Public Health

**Length**: 9:42

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=Jd3gFT0-C4s")
```

Link to video transcript [here](chapter4.1_transcript.qmd).

------------------------------------------------------------------------

## 4.2 Confounding, Chance, and Bias {#sec-section2}

**Title**: Confounding, chance, and bias

**Description**: This video provides an overview over three types of error that can lead to misleading or incorrect results in medical studies: confounding, chance, and bias.

**Presenter(s)**: Cochrane Austria, Department for Evidence-based Medicine and Evaluation, Danube University Krems

**Length**: 9:50

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=bcfg9kcxeuU")
```

Link to video transcript [here](chapter4.2_transcript.qmd).

------------------------------------------------------------------------

## 4.3 Genetic study designs {#sec-section3}

### Twin studies

**Title**: What are "Twin studies"?

**Description**: In this animation, we explore how studying many pairs of twins allows us to estimate how much a trait comes from our genes and how much is influenced by the environment.

**Presenter(s)**: OpenLearn from the Open University

**Length**: 2:13

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=BTYCv1ObZrI&t=1s")
```

Link to video transcript [here](chapter4.3_transcript.qmd#sec-video1).

------------------------------------------------------------------------

### Choosing the Right Study Design in Genetic Epidemiology

**Title**: Choosing the Right Study Design in Genetic Epidemiology

**Description**:

**Presenter(s)**: Duncan Thomas, University of Southern California

**Length**: 34:31

Link to video [here](https://hstalks.com/t/25/choosing-the-right-study-design-in-genetic-epidemi/).

Link to video transcript [here](chapter4.3_transcript.qmd#sec-video2).
