---
title: "Chapter 10: Other Considerations"
---

Chapter goals:

# 10.1 A Career in Psychiatric Genetics {#sec-section1}

**Title**: How does Genetics Affect our Mental Health?

**Presenter(s)**:

-   Cathryn Lewis, PhD (Social, Genetic, and Developmental Psychiatry Centre, King's College London)

-   Alex Curmi, MD (Maudsley NHS Foundation Trust)

**Level**: Beginner

**Length**: 58:54

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=8jtoLRmC5F4")
```

Link to video transcript [here](chapter10.1_transcript.qmd).

------------------------------------------------------------------------

# 10.2 Caution in Genetic Prediction {#sec-section2}

**Title**: Predicting the likelihood of future psychiatric disorders: a closer look, and some cautions.

**Presenter(s)**: Howard Edenberg, PhD (Department of Medical and Molecular Genetics, Indiana University School of Medicine)

**Level**: Beginner

**Length**: 4:30

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=P8Odk5HuY50")
```

Link to video transcript [here](chapter10.2_transcript.qmd).

------------------------------------------------------------------------

# 10.3 Small Effect Sizes {#sec-section3}

**Title**: Small effect sizes

**Presenter(s)**: Howard Edenberg, PhD (Department of Medical and Molecular Genetics, Indiana University School of Medicine)

**Level**: Beginner

**Length**: 4:05

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=l5tsUyhXjt4")
```

Link to video transcript [here](chapter10.3_transcript.qmd).

------------------------------------------------------------------------

# 10.4 Language in Genetics Research {#sec-section4}

**Content coming soon!**

**Presenter(s)**:

-   Carina Seah, MS (Department of Genetics and Genomic Sciences, Icahn School of Medicine at Mount Sinai)

-   Kayla Townsley, MS (Department of Genetics and Genomic Sciences, Icahn School of Medicine at Mount Sinai)

**Level**: Beginner

------------------------------------------------------------------------

# 10.5 Equitable Collaboration for LMIC Researchers {#sec-section5}

**Content coming soon!**

------------------------------------------------------------------------

# 10.6 GDPR for Dummies: A Survival Guide for Genetics Research {#sec-section6}

**Title**: GDPR for dummies: A survival guide for genetics research

**Presenter(s)**: Heidi Beate Bentzen, LLM (Centre for Medical Ethics, University of Oslo)

**Level**: Beginner

**Length**: 23:46

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=bxCV2Z5i_yI&t=8s")
```

Link to video transcript [here](chapter10.6_transcript.qmd).
