---
title: "Chapter 7: Ancestry-Specific Analyses and Considerations"
---

> *Chapter goals:*
>
> 1.  *Understand that ancestry is an important consideration for performing genetic analyses*
>
> 2.  *Understand how population stratification can affect GWAS results*
>
> 3.  *Understand considerations for PRS across different ancestries and how to calculate cross-ancestry PRS*
>
> 4.  *Understand cross-ancestry methods for GWAS and post-GWAS analyses*

# Chapter 7.1: Cross-ancestry analysis {#sec-section1}

**Title**: Cross-ancestry PTSD and Polygenic Findings, and the Cross-Population SIG of the PGC

**Presenter(s)**: Laramie Duncan, PhD (Stanford University)

**Level**: Beginner friendly

**Length**: Video starts at 00:00 and goes until 13:39 (13:39 total)

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=p3m6MM5b-UU&t=148s")
```

Link to video transcript [here](chapter7_transcript.qmd#sec-video1).

\[Note: This paper is now available as [*Genome-wide Association Studies in Ancestrally Diverse Populations: Opportunities, Methods, Pitfalls, and Recommendations: https://doi.org/10.1016/j.cell.2019.08.051*\]](https://doi.org/10.1016/j.cell.2019.08.051)

------------------------------------------------------------------------

# Chapter 7.2: Ancestry-specific PRS {#sec-section2}

**Title**: Clinical use of current polygenic risk scores may exacerbate health disparities

**Presenter(s)**: Alicia Martin, PhD (Broad Institute)

**Level**: Beginner friendly

**Length**: Video starts at 32:39 and goes to 47:06 (15:45 total)

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=p3m6MM5b-UU&t=2321s")
```

Link to video transcript [here](chapter7_transcript.qmd#sec-video2).

------------------------------------------------------------------------

# Chapter 7.3: Local Ancestry and Admixed Populations {#sec-section3}

**Title**: Tractor: A framework enabling the well-calibrated genomic analysis of psychiatric traits across admixed populations

**Presenter(s)**: Elizabeth Atkinson, PhD (Department of Molecular and Human Genetics, Baylor College of Medicine)

**Level**: Beginner/Intermediate

**Length**: Video starts at 14:50 and ends at 32:38 (total 19:28)

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=p3m6MM5b-UU&t=890s")
```

Link to video transcript [here](chapter7_transcript.qmd#sec-video3).

Link to Tractor [GitHub](https://github.com/Atkinson-Lab/Tractor) and [Tutorial](https://github.com/Atkinson-Lab/Tractor-tutorial).
