---
title: "SNP Heritability and Genetic Correlation"
---

# LAVA

**Title**: LAVA Tutorial

**Presenter(s)**: Josefin Werme

**Level**: Intermediate

**Length**: 16:08

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=ovfYfcVLLO0")
```

Link to video transcript [here](software_correlation_transcript.qmd#sec-video1).

------------------------------------------------------------------------

# GCTA-GREML

**Title**: Estimating SNP-based heritability with GCTA-GREML

**Presenter(s)**: Jian Yang

**Level**: Intermediate

**Length**: 18:28

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=kk9Fkg0trT8")
```

Link to video transcript [here](software_correlation_transcript.qmd#sec-video2).

------------------------------------------------------------------------

# Genetic Architecture

**Title**: Tracking 12 years of genetic architecture estimates of schizophrenia

**Presenter(s)**: Naomi Wray

**Level**: Beginner

**Length**: 15:26

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=ze6Oe5_DwWU")
```

Link to video transcript [here](software_correlation_transcript.qmd#sec-video3).
