---
title: "Datasets"
---

> Large genetic and genomic datasets are becoming more publicly available for research use. In the following videos, Drs. Palotie, Liu, and O'Donnell-Luria give introductions to large datasets available from the FinnGen, PsychENCODE, and gnomAD resources respectively.
>
> FinnGen is a large consortia of both public and private institutions (universities, hospitals, biobanks, pharmaceutical companies, etc.) to collect and analyse data from over 500,000 Finnish biobank participants, with the ultimate goal of finding novel insight into prevention, treatment, and diagnosis of disease.
>
> The PsychENCODE consortium is composed of many academic institutions whose goal is the generation of large-scale genomic data (gene expression and regulation) from individuals with psychiatric disorders across different brain regions. This includes generation of both bulk and single-cell RNA sequencing, ChIPseq, Hi-C, ATACseq, and other measures of chromatin activity for multiple brain regions across different developmental stages.
>
> The **G**e**nom**e **A**ggregation **D**atabase, or gnomAD, is a large reference of human genetic variation from whole exome data.

# FinnGen {#sec-section1}

**Title**: Intro to FinnGen

**Presenter(s)**: Aarno Palotie, MD, PhD (Institute for Molecular Medicine Finland (FIMM), University of Helsinki)

**Level**: Beginner

**Length**: 19:24

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=WPprJg7fIBg")
```

Link to video transcript [here](software_datasets_transcript.qmd#sec-video1).

Link to [FinnGen website](https://www.finngen.fi/en).

------------------------------------------------------------------------

# PsychENCODE {#sec-section2}

**Title**: Introduction to PsychENCODE

**Presenter(s)**: Chunyu Liu, PhD (Department of Psychiatry, SUNY Upstate Medical University)

**Level**: Beginner

**Length**: 19:34

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=mnnnmuk09JE")
```

Link to video transcript [here](software_datasets_transcript.qmd#sec-video2).

Link to [PsychENCODE website](https://psychencode.synapse.org) and [PsychSCREEN website](http://psychscreen.beta.wenglab.org/main/search).

------------------------------------------------------------------------

# GnomAD

**Title**: gnomAD: Using large genomic data sets to interpret human genetic variation

**Presenter(s)**: Anne O'Donnell-Luria, PhD (Broad Institute of Harvard and MIT)

**Level**: Beginner

**Length**: 54:20

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=J29_Pf8Ti1s")
```

Link to video transcript [here](software_datasets_transcript.qmd#sec-section3).

Link to [gnomAD website](https://gnomad.broadinstitute.org/).
