---
title: "GWAS"
---

# Genome-Wide Association Studies in PLINK {#sec-section1}

**Title**: Genome-wide Association Studies

**Presenter(s)**: Hailiang Huang, PhD (Department of Medicine, Harvard Medical School)

**Level**: Intermediate

**Length**: 15:47

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=z7QIRbXcbDw")
```

Link to video transcript [here](software_gwas_transcript.qmd#sec-video1).

[Link to tutorial GitHub and datasets](https://github.com/thehuanglab/gwas-tutorial).

------------------------------------------------------------------------

# Genotype QC {#sec-section2}

**Title**: How to run quality control on genome-wide genotyping data

**Presenter(s)**: Jonathan Coleman, PhD (Social, Genetic, and Developmental Psychiatry Centre, King's College London)

**Level**: Intermediate

**Length**: 16:19

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=koHuMAeWQ5E")
```

Link to video transcript [here](software_gwas_transcript.qmd#sec-video2).

Link to [PGC RICOPILI pipeline](https://sites.google.com/a/broadinstitute.org/ricopili/).

Link to [Coleman lab GitHub](https://github.com/ColemanResearchGroup/gwas_scripts) GWAS scripts.

------------------------------------------------------------------------

# Tractor: GWAS with Admixed Individuals {#sec-section3}

**Title**: Tractor: Enabling GWAS in admixed cohorts

**Presenter(s)**: Elizabeth Atkinson, PhD (Department of Molecular and Human Genetics, Baylor College of Medicine)

**Level**: Intermediate

**Length**: 18:38

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=UUviC0u54rE")
```

Link to video transcript [here](software_gwas_transcript.qmd#sec-video3).

[Link to tutorial GitHub and datasets](https://atkinson-lab.github.io/Tractor-tutorial/).

------------------------------------------------------------------------

# Genotype Calling and Imputation with MoChA {#sec-section4}

**Title**: Genotype calling/imputation on the cloud: MoChA pipleine

**Presenter(s)**: Giulio Genovese, PhD (Program in Medical and Population Genetics, Broad Institute of Harvard and MIT)

**Level**: Intermediate

**Length**: 15:13

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=QWECuWknTVY")
```

Link to video transcript [here](software_gwas_transcript.qmd#sec-video4).

Link to [MoChA GitHub](https://github.com/freeseek/mocha) and [Webpage](https://software.broadinstitute.org/software/mocha/).

------------------------------------------------------------------------

# SAIGE: Family-Based GWAS {#sec-section5}

**Title**: Genetic association studies in large-scale biobanks and cohorts

**Presenter(s)**: Wei Zhou, PhD (Analytic and Translational Genetics Unit, The Broad Institute of Harvard and MIT)

**Level**: Intermediate

**Length**: 27:17

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=ZrWFOSAdc74")
```

Link to video transcript [here](software_gwas_transcript.qmd#sec-video5).

Link to GitHub and Wiki Tutorials for [SAIGE](https://github.com/weizhouUMICH/SAIGE/wiki/SAIGE-Hands-On-Practical---09-2021-updated) and [SAIGE-GENE](https://github.com/weizhouUMICH/SAIGE/wiki/SAIGE-GENE:-Hands-On-Practice-(09-2021-updated)).

------------------------------------------------------------------------

# CC-GWAS {#sec-section6}

**Title**: Tutorial to use CC-GWAS to identify loci with different allele frequency among cases of different disorders.

**Presenter(s)**: Wouter Peyrot, MD, PhD (Department of Psychiatry, Amsterdam UMC)

**Level**: Intermediate

**Length**: 21:46

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=Hr1upC0U8Q8")
```

Link to video transcript [here](software_gwas_transcript.qmd#sec-video6).

[Link to tutorial GitHub and datasets](https://github.com/wouterpeyrot/CCGWAS).
