---
title: "PRS"
---

------------------------------------------------------------------------

# Comparison of PRS methods

**Title**: Comparison of PRS methods

**Description**:

**Presenter(s)**: Guiyan Ni, Program in Complex Trait Genomics, Institute for Molecular Bioscience, University of Queensland

**Level**: Intermediate

**Length**: 13:52

```{r, echo=FALSE}
library("vembedr")
embed_url("https://youtu.be/H18i2Oe05Gw")
```

Link to video transcript [here](software_prs_transcript.qmd#sec-video1).

------------------------------------------------------------------------

# PRS-CS and PRS-CSx

## PRS-CSx to perform PRS analysis

**Title**: Hands-on tutorial of using PRS-CSx to perform multi-ancestry PRS analysis

**Description**:

**Presenter(s)**: Tian Ge, Yunfeng Ruan, Stanley Center, Broad Institute

**Level**: Intermediate

**Length**: 8:28

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=ERAuCaoJ4Wk")
```

Link to video transcript [here](software_prs_transcript.qmd#sec-video2).

## PRS in Ancestrally-diverse Populations

**Title**: PRS-CSx: Improving Cross-Population Polygenic Prediction using Coupled Continuous Shrinkage Priors

**Presenter(s)**: Tian Ge

**Level**: Intermediate

**Length**: 56:46

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=UM554Hi_v-Y")
```

Link to video transcript [here](software_prs_transcript.qmd#sec-video3).

------------------------------------------------------------------------

# LDpred2

**Title**: How to run LDpred2

**Description**:

**Presenter(s)**: Florian Privé

**Level**: Intermediate

**Length**: 15:00

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=aya8WsNAu6U")
```

Link to video transcript [here](software_prs_transcript.qmd#sec-video5).

Link to tutorial test datasets [here](https://privefl.github.io/bigsnpr/articles/LDpred2.html).

------------------------------------------------------------------------

# PRSice2 and LASSOSUM

**Title**: PRSice: Basic Polygenic analysis

**Description**:

**Presenter(s)**: Shing Wan Choi

**Level**: Intermediate

**Length**: 25:35

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=Q-hRBQ7kAPM")
```

Link to video transcript [here](software_prs_transcript.qmd#sec-video4).

------------------------------------------------------------------------

# SBayesR

**Title**: How to run SBayesR

**Description**:

**Presenter(s)**: Jian Zeng

**Level**: Intermediate

**Length**: 8:09

```{r, echo=FALSE}
library("vembedr")
embed_url("https://youtu.be/nUNU7TZMui4")
```

Link to video transcript [here](software_prs_transcript.qmd#sec-video6).

------------------------------------------------------------------------

# Vertical Transmission and Genetic Nurture with PRS

**Title**: Estimating Vertical Transmission and Genetic Nurture using Polygenic Scores

**Description**:

**Presenter(s)**: Jared Balbona, Yongkang Kim, Matthew Keller

**Level**: Intermediate

**Length**: 21:30

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=w7wB0HsGAJE")
```

Link to video transcript [here](software_prs_transcript.qmd#sec-video7).
