---
title: "GWAS"
---

------------------------------------------------------------------------

# Genome-wide Association Studies in PLINK

**Title**: Genome-wide Association Studies

**Presenter(s)**: Hailiang Huang

**Level**: Intermediate

**Length**: 15:47

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=z7QIRbXcbDw")
```

Link to video transcript [here](software_gwas_transcript.qmd#sec-video1).

[Link to tutorial GitHub and datasets](https://github.com/thehuanglab/gwas-tutorial).

------------------------------------------------------------------------

# Genotype QC

**Title**: How to run quality control on genome-wide genotyping data

**Presenter(s)**: Joni Coleman

**Level**: Intermediate

**Length**: 16:19

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=koHuMAeWQ5E")
```

Link to video transcript [here](software_gwas_transcript.qmd#sec-video2).

Link to [PGC RICOPILI pipeline](https://sites.google.com/a/broadinstitute.org/ricopili/).

Link to [Coleman lab GitHub](https://github.com/ColemanResearchGroup/gwas_scripts) GWAS scripts.

------------------------------------------------------------------------

# Tractor: GWAS with admixed individuals

**Title**: Tractor: Enabling GWAS in admixed cohorts

**Presenter(s)**: Elizabeth Atkinson

**Level**: Intermediate

**Length**: 18:38

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=UUviC0u54rE")
```

Link to video transcript [here](software_gwas_transcript.qmd#sec-video3).

[Link to tutorial GitHub and datasets](https://atkinson-lab.github.io/Tractor-tutorial/).

------------------------------------------------------------------------

# Genotype Calling and Imputation with MoChA

**Title**: Genotype calling/imputation on the cloud: MoChA pipleine

**Presenter(s)**: Giulio Genovese

**Level**: Intermediate

**Length**: 15:13

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=QWECuWknTVY")
```

Link to video transcript [here](software_gwas_transcript.qmd#sec-video4).

Link to [MoChA GitHub](https://github.com/freeseek/mocha) and [Webpage](https://software.broadinstitute.org/software/mocha/).

------------------------------------------------------------------------

# SAIGE: Family-based GWAS

**Title**: Genetic association studies in large-scale biobanks and cohorts

**Presenter(s)**: Wei Zhou

**Level**: Intermediate

**Length**: 27:17

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=ZrWFOSAdc74")
```

Link to video transcript [here](software_gwas_transcript.qmd#sec-video5).

Link to GitHub and Wiki Tutorials for [SAIGE](https://github.com/weizhouUMICH/SAIGE/wiki/SAIGE-Hands-On-Practical---09-2021-updated) and [SAIGE-GENE](https://github.com/weizhouUMICH/SAIGE/wiki/SAIGE-GENE:-Hands-On-Practice-(09-2021-updated)).

------------------------------------------------------------------------

# CC-GWAS

**Title**: Tutorial to use CC-GWAS to identify loci with different allele frequency among cases of different disorders.

**Presenter(s)**: Wouter Peyrot

**Level**: Intermediate

**Length**: 21:46

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=Hr1upC0U8Q8")
```

Link to video transcript [here](software_gwas_transcript.qmd#sec-video6).

[Link to tutorial GitHub and datasets](https://github.com/wouterpeyrot/CCGWAS).
