---
title: "PRS"
---

------------------------------------------------------------------------

# Comparison of PRS Methods {#sec-section1}

**Title**: Comparison of PRS methods

**Presenter(s)**: Guiyan Ni, PhD (Institute for Molecular Bioscience, University of Queensland)

**Level**: Intermediate

**Length**: 13:52

```{r, echo=FALSE}
library("vembedr")
embed_url("https://youtu.be/H18i2Oe05Gw")
```

Link to video transcript [here](software_prs_transcript.qmd#sec-video1).

------------------------------------------------------------------------

# PRS-CS and PRS-CSx {#sec-section2}

## PRS-CSx to perform PRS analysis

**Title**: Hands-on tutorial of using PRS-CSx to perform multi-ancestry PRS analysis

**Presenter(s)**:

-   Tian Ge, PhD (Center for Genomic Medicine, Mass General Research Institute)

-   Yunfeng Ruan, PhD (Stanley Center for Psychiatric Research, Broad Institute of MIT and Harvard)

**Level**: Intermediate

**Length**: 8:28

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=ERAuCaoJ4Wk")
```

Link to video transcript [here](software_prs_transcript.qmd#sec-video2).

------------------------------------------------------------------------

## PRS in Ancestrally-Diverse Populations

**Title**: PRS-CSx: Improving Cross-Population Polygenic Prediction using Coupled Continuous Shrinkage Priors

**Presenter(s)**: Tian Ge, PhD (Center for Genomic Medicine, Mass General Research Institute)

**Level**: Intermediate

**Length**: 56:46

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=UM554Hi_v-Y")
```

Link to video transcript [here](software_prs_transcript.qmd#sec-video3).

------------------------------------------------------------------------

# LDpred2 {#sec-section3}

**Title**: How to run LDpred2

**Presenter(s)**: Florian Privé, PhD (National Centre for Register-Based Research, Aarhus University)

**Level**: Intermediate

**Length**: 15:00

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=aya8WsNAu6U")
```

Link to video transcript [here](software_prs_transcript.qmd#sec-video5).

Link to tutorial test datasets [here](https://privefl.github.io/bigsnpr/articles/LDpred2.html).

------------------------------------------------------------------------

# PRSice2 and LASSOSUM {#sec-section4}

**Title**: PRSice: Basic Polygenic analysis

**Presenter(s)**: Shing Wan Choi, PhD (Regeneron)

**Level**: Intermediate

**Length**: 25:35

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=Q-hRBQ7kAPM")
```

Link to video transcript [here](software_prs_transcript.qmd#sec-video4).

------------------------------------------------------------------------

# SBayesR {#sec-section5}

> *\[Note\]: At the moment the video input does not have sound. We are working on getting a working version up very soon!*

**Title**: How to run SBayesR

**Presenter(s)**: Jian Zeng, PhD (Institute for Molecular Bioscience, University of Queensland)

**Level**: Intermediate

**Length**: 8:09

```{r, echo=FALSE}
library("vembedr")
embed_url("https://youtu.be/nUNU7TZMui4")
```

Link to video transcript [here](software_prs_transcript.qmd#sec-video6).

------------------------------------------------------------------------

# Vertical Transmission and Genetic Nurture with PRS {#sec-section6}

**Title**: Estimating Vertical Transmission and Genetic Nurture using Polygenic Scores

**Presenter(s)**: Jared Balbona, PhD (Institute for Behavioral Genetics, University of Colorado Boulder)

**Coauthor(s)**:

-   Yongkang Kim, PhD (Janssen Pharmaceutical Companies of Johnson & Johnson)

-   Matthew Keller, PhD (Institute for Behavioral Genetics, University of Colorado Boulder)

**Level**: Intermediate

**Length**: 21:30

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=w7wB0HsGAJE")
```

Link to video transcript [here](software_prs_transcript.qmd#sec-video7).

Link to [code and practice data on Google drive](https://drive.google.com/drive/folders/1C7LrfLJO8cbhwGHIApc4yc2Or-skcZ1E).

Link to [OpenMx R package](https://openmx.ssri.psu.edu/).
