---
title: "Gene Set Identification"
---

------------------------------------------------------------------------

# MAGMA {#sec-section1}

**Title**: Gene- and gene-set analysis in MAGMA

**Presenter(s)**: Christiaan de Leeuw, PhD (Department of Complex Trait Genetics, Vrije Universiteit Amsterdam)

**Level**: Intermediate

**Length**: 12:09

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=nOxHTYz-z7s")
```

Link to video transcript [here](software_geneset_transcript.qmd#sec-video1).

------------------------------------------------------------------------

# H-MAGMA {#sec-section2}

**Title**: Annotating Genetic Variants to Target Genes Using H-MAGMA

**Presenter(s)**: Nancy Y.A. Sey, PhD (National Human Genome Research Institute)

**Level**: Intermediate

**Length**: 10:09

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=mhQ5WQ2HrUc")
```

Link to video transcript [here](software_geneset_transcript.qmd#sec-video2).

[Link to GitHub tutorial and datasets](https://github.com/thewonlab/H-MAGMA).

------------------------------------------------------------------------

# E-MAGMA {#sec-section3}

**Title**: E-MAGMA: an eQTL-informed method to identify risk genes using genome-wide association study summary statistics

**Presenter(s)**:

-   Zac Gerring, PhD (Translational Neurogenomics, QIMR Berghofer Medical Research Institute)

-   Eske Derks, PhD (Translational Neurogenomics, QIMR Berghofer Medical Research Institute)

**Level**: Intermediate

**Length**: 7:16

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=NEk-h070AbU")
```

Link to video transcript [here](software_geneset_transcript.qmd#sec-video3).

[Link to GitHub tutorial and datasets](https://github.com/eskederks/eMAGMA-tutorial).

------------------------------------------------------------------------

# PRSet {#sec-section4}

**Title**: How to run pathway specific Polygenic Risk Scores

**Presenter(s)**: Judit García-González, PhD (Department of Genetics and Genomic Sciences, Icahn School of Medicine at Mount Sinai)

**Level**: Intermediate

**Length**: 15:28

```{r, echo=FALSE}
library("vembedr")
embed_url("https://www.youtube.com/watch?v=IrriXu26h_E")
```

Link to video transcript [here](software_geneset_transcript.qmd#sec-video4).

Link to [PRSet website](https://choishingwan.github.io/PRSice/quick_start_prset/) (part of the PRSice-2 website).
